Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-06-16 eventPointTwoLoop.tex
small changes
fixed dead section label, changed formatting, re-aranged some citations

\documentclass[12pt,letter]{article}
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scalebox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scalebox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

\usepackage{tikz-feynman} 
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={funRed},
    citecolor={funBlue},
    urlcolor={funBlue}
}
\tikzset{graviton/.style={decorate, decoration={snake, amplitude=.4mm, segment length=1.5mm, pre length=.5mm, post length=.5mm}, double}}
\tikzfeynmanset{compat=1.1.0}
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
          %% Load the `external` library
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
  prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
  },
]

@@

\definecolor{funRed}{RGB}{139,1,0}
\definecolor{funBlue}{RGB}{1,76,125}
\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgreen}{RGB}{25,176,146}
\definecolor{hgreen1}{RGB}{175,230,175}
\definecolor{hblue}{RGB}{52,152,219}
\definecolor{hbluedark}{RGB}{36, 106, 160}
\definecolor{hblue1}{RGB}{255,255,166}
\definecolor{hred}{RGB}{139,1,0}
\definecolor{hreddark}{RGB}{151, 58, 81}
\definecolor{hred1}{RGB}{255,155,155}
\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgrey4}{RGB}{75,75,75}
\definecolor{hgrey5}{RGB}{50,50,50}
\definecolor{hgrey3}{RGB}{100,100,100}
\definecolor{hgrey}{RGB}{125,125,125}
\definecolor{hgrey2}{RGB}{125,125,125}
\definecolor{hgrey1}{RGB}{150,150,150}
\definecolor{hgrey0}{RGB}{175,175,175}
\definecolor{darkgreen}{RGB}{59,126,108}

@@

\newcommand{\scaleIntAfermion}[4]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[fermion, ultra thick,out=60,in=120,min distance=0.4cm,hred](mid3),
(mid3) --[fermion, ultra thick,out=-120,in=-60,min distance=0.4cm,hred](mid1),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\scaleIntAScalar}[4]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[ ultra thick,out=60,in=120,min distance=0.4cm,hblue](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm,hblue](mid3),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\newcommand{\scaleIntAvector}[4]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[photon, ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[photon, ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\newcommand{\scaleIntBtune}[8]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-1.4,.5);
\vertex (a2) at (1.2,1);
\vertex [dot,scale=3](mid1) at (0.7,0.5){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0.7,0.5){};
\vertex [dot,scale=3](mid3) at (-0.7,0.5){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (-0.7,0.5){};
\vertex [dot,scale=3](mid5) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid6) at (0,0){};
\vertex (a3) at (-.5, -.5);
\vertex (a4) at (.5, -.5);
\diagram{
(mid3) --[photon, ultra thick](a1),
(mid1) --[photon, ultra thick](a2),
(mid1) --[#1, ultra thick,out=120,in=60,min distance=0.1cm,#5](mid3),
(mid1) --[#2, ultra thick,#6](mid3),

@@

\newcommand{\scaleIntCtune}[4]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {};
\vertex (a2) at (-.8, 0.6) {};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex [dot,scale=3](mid5) at (2,0){};
\vertex [dot,scale=2.5,hgrey0](mid6) at (2,0){};
\vertex (a3) at (2.8, .6) {};
\vertex (a4) at (2.8, -.6) {};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[#1, ultra thick,out=60,in=120,min distance=0.4cm,#2](mid3),
(mid3) --[#1, ultra thick,in=-60,out=-120,min distance=0.4cm,#2](mid1),
(mid3) --[#3, ultra thick,out=60,in=120,min distance=0.4cm,#4](mid5),
(mid5) --[#3, ultra thick,in=-60,out=-120,min distance=0.4cm,#4](mid3),
(mid5) --[photon, ultra thick](a4),
(mid5) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\title{Insights of Even-point Multi-loop Unitarity: Exponentiation, Anomalies and Evanescence}

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  For the first time, we exploit special properties of these even-point theories to generalize the direct extraction of Feynman integral coefficients using $D$-dimensional unitarity cuts to the multi-loop case. In doing so, we find that the leading IR divergence of NLSM amplitudes exponentiates. We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms required to restore $U(1)$ behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  We also demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator in the $D$-dimensional description of Born-Infeld theory. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violates $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order. This is shown by reproducing our unitarity based result with a loop-level double copy between $\mathcal{N}=4$ super-Yang-Mills numerators and our two-loop NLSM amplitudes. This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

\section{Introduction}\label{sec:intro}
Recent decades have witnessed a renaissance in our ability to compute scattering amplitudes to high orders in perturbation theory. At the epicenter of this explosion in $S$-matrix data is the unitarity method \cite{UnitarityMethod, Fusing, BDKUniarityReview}, which bypasses the standard Feynman diagram approach by constructing amplitudes directly from on-shell information. At one-loop, the unitarity method allows one to extract amplitudes by directly computing a series of unitarity cuts \cite{Forde:2007mi}. On-shell methods have furthermore unveiled novel amplitude-level structure, like the color-kinematics duality \cite{BCJ} and the double-copy construction \cite{BCJLoop}, which are completely obscured by a Lagrangian-level description of the physics. While much of the research in perturbative calculations thus far has focused on gauge theory and gravity, recent literature has investigated to what extent the $S$-matrix of effective field theory is constrained by on-shell data \cite{Adams:2006sv,Cheung:2015ota,Cheung:2016drk,Carrasco:2019yyn,Arkani-Hamed:2020blm,Carrasco:2021ptp,Chi:2021mio,Bonnefoy:2021qgu,Carrasco:2022lbm,Carrasco:2022sck,Green:2022slj,Pavao:2022kog,Chen:2022shl,Chen:2023dcx,Brown:2023srz}. In this work, we take a quantum leap to the multi-loop sector of effective field theory. As we embark on this journey, we anticipate that further probes of perturbative effective field theory data will unravel a rich tapestry of discoveries, adding to the current knowledge base and paving the way for future advancements in the field.

@@

By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Adams:2006sv,Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2015ota,Cheung:2016drk,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed} . In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 

@@

\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu - \varphi \partial_\mu \varphi)

@@

The action above in \eqn{DBIVALag} can be similarly re-expressed in terms of our canonical spacetime coordinates, giving us the following flat-space Lagrangian for maximally supersymmetric Born-Infeld theory in $D=10$: 
\begin{equation}
 \label{dbivaLag}
    \alpha'^{\,2}\mathcal{L}^{\text{DBIVA}}s = \sqrt{\text{det} \left(\eta_{\mu\nu}- \alpha' F_{\mu\nu} + \alpha'^{\,2} (\bar{\lambda} \Gamma_\mu \partial_\nu \lambda)- \alpha'^4(\bar{\lambda} \Gamma^\rho \partial_\mu \lambda )(\bar{\lambda} \Gamma_\rho \partial_\nu \lambda) \right)}
  \,.
\end{equation}
To recover the $D=4$ theory, we simply dimensionally reduce in a way that preserves the maximum number of supercharges, which is $\mathcal{N}=4$ in $D=4$. While this is obscured at the level of the Lagrangian, implementing a dimensional reduction that preserves the supercharges of \eqn{dbivaLag} is rather straight forward in an on-shell framework. 
\be
\mathcal{M}^{\text{DBIVA}} = A^{\text{NLSM}} \otimes A^{\text{sYM}}
\ee
Since the double copy construction holds $D$-dimensionally at tree-level \cite{BCJreview}, double-copying maximal sYM with NLSM will yield the analogous spectrum of maximally supersymmetry DBIVA theory, whether in $D=10$, as described above, or in $D=4$, as we will study in this work. 
\paragraph{Abelian Open String}
Similar to the amplitudes of DBIVA theories described above, the tree-level amplitudes for the open superstring (OSS) ~\cite{Mafra:2011nv,Mafra:2011nw} likewise permit a double copy construction \cite{Broedel:2013tta}. To construct open-superstring amplitudes we simply double copy maximally supersymmetric super-Yang-Mills (sYM) and $Z$-theory amplitudes~\cite{Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc}, as follows,
\begin{equation}
A^{\text{OSS}}= Z \otimes A^{ \text{sYM}}
\end{equation}
where $\otimes$ implements the field theory double copy \cite{KLT, BCJ} described above. Moreover, the amplitudes of DBIVA pop out of the above construction as the field theory limit of the \textit{abelian open superstring} \cite{Green:1982sw}. To recover the abelian sector, one simply sums over all color orderings on the side of the bi-colored $Z$-theory amplitudes that obey string monodromy relations. In the field theory limit, the observables for so-called \textit{abelian} $Z$-theory are simply the amplitudes generated by the NLSM Lagrangian \cite{Carrasco:2016ldy}. That is, given a $n$-point amplitude in $Z$-theory, with field theory ordering $a=(a_1,a_2,...,a_n)$ and string theoretic color ordering $A=(A_1,A_2,...,A_n)$, one finds the following relation:
\begin{equation}
A^{\text{NLSM}}{(a_1,a_2,...,a_n)} \equiv \lim_{\alpha' \to 0} (\alpha')^{2-n}  Z_{\times}{(a_1,a_2,...,a_n)}
\end{equation}
where $Z_{\times}$ are abelian $Z$-theory amplitudes of \cite{Carrasco:2016ldy}, defined by summing all possible orderings of Chan-Paton factors:
\begin{equation}
Z_{\times}(a_1,a_2,...,a_n)\equiv \sum_{A \in S^{n-1}} Z_{(A_1,...,A_{n-1},n)}{(a_1,a_2,...,a_n)}
\end{equation}
Since the field theory limit of abelian $Z$-theory produces NLSM amplitudes, the field theory limit of the abelian OSS, gives rise to DBIVA observables at leading order:
\begin{equation}
\boxed
{\begin{aligned} 
\lim_{\alpha' \to 0 }Z^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{\text{BAS}}\equiv A^{\text{NLSM}}
\\
\Rightarrow \quad\lim_{\alpha' \to 0 }A^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{ \text{sYM}}\equiv A^{\text{DBIVA}}
\end{aligned}}
\end{equation}
Additional details on the bi-colored $Z$-theory amplitudes these string-theoretic double-copy constructions can be found in \cite{Broedel:2013tta,Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc,Azevedo:2018dgo} and references therein. We have boxed the two classes of amplitudes above, NLSM and DBIVA, to emphasize that these are the two theories that we will focus on in this work. However, the methods that we develop in the next section apply much more broadly to the full catalog of even-point theories. One could

@@

@@

\boxed{

@@

}

@@

@@

where the $D$-dimensional charge conjugation operator can be defined in terms of the gamma matrices, $\Gamma_\mu$ as follows:

@@

\boxed{\begin{aligned}

@@

\end{aligned}}

>>>>

\documentclass[11pt,letter]{article}
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scalebox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scalebox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

\usepackage{tikz-feynman} 
\usepackage[framemethod=default]{mdframed}
\newmdenv[skipabove=7pt,
skipbelow=7pt,
rightline=false,
leftline=false,
topline=false,
bottomline=false,
backgroundcolor=gray!10,
linecolor=gray,
innerleftmargin=5pt,
innerrightmargin=5pt,
innertopmargin=5pt,
innerbottommargin=5pt,
leftmargin=0cm,
rightmargin=0cm,
linewidth=4pt]{eBox}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={funRed},
    citecolor={funBlue},
    urlcolor={funBlue}
}
\tikzset{graviton/.style={decorate, decoration={snake, amplitude=.4mm, segment length=1.5mm, pre length=.5mm, post length=.5mm}, double}}
\tikzfeynmanset{compat=1.1.0}
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
          %% Load the `external` library
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
  prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
  },
]

@@

\definecolor{funRed}{RGB}{139,1,0}
\definecolor{funBlue}{RGB}{1,76,125}
\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgreen}{RGB}{25,176,146}
\definecolor{hgreen1}{RGB}{175,230,175}
\definecolor{hblue}{RGB}{52,152,219}
\definecolor{hbluedark}{RGB}{36, 106, 160}
\definecolor{hblue1}{RGB}{255,255,166}
\definecolor{hred}{RGB}{139,1,0}
\definecolor{hreddark}{RGB}{151, 58, 81}
\definecolor{hred1}{RGB}{255,155,155}
\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgrey4}{RGB}{75,75,75}
\definecolor{hgrey5}{RGB}{50,50,50}
\definecolor{hgrey3}{RGB}{100,100,100}
\definecolor{hgrey}{RGB}{125,125,125}
\definecolor{hgrey2}{RGB}{125,125,125}
\definecolor{hgrey1}{RGB}{150,150,150}
\definecolor{hgrey0}{RGB}{175,175,175}
\definecolor{hgreyLight}{RGB}{242,242,242}
\definecolor{darkgreen}{RGB}{59,126,108}

@@

\newcommand{\scaleIntAfermion}[4]{ {
\begin{tikzpicture}[baseline=-3]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[fermion, ultra thick,out=60,in=120,min distance=0.4cm,hred](mid3),
(mid3) --[fermion, ultra thick,out=-120,in=-60,min distance=0.4cm,hred](mid1),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\scaleIntAScalar}[4]{ {
\begin{tikzpicture}[baseline=-3]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[ ultra thick,out=60,in=120,min distance=0.4cm,hblue](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm,hblue](mid3),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\newcommand{\scaleIntAvector}[4]{ {
\begin{tikzpicture}[baseline=-3]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[photon, ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[photon, ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\newcommand{\scaleIntBtune}[8]{ {
\begin{tikzpicture}[baseline=4]
\begin{feynman}
\vertex (a1) at (-1.4,.5);
\vertex (a2) at (1.2,1);
\vertex [dot,scale=3](mid1) at (0.7,0.5){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0.7,0.5){};
\vertex [dot,scale=3](mid3) at (-0.7,0.5){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (-0.7,0.5){};
\vertex [dot,scale=3](mid5) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid6) at (0,0){};
\vertex (a3) at (-.5, -.5);
\vertex (a4) at (.5, -.5);
\diagram{
(mid3) --[photon, ultra thick](a1),
(mid1) --[photon, ultra thick](a2),
(mid1) --[#1, ultra thick,out=120,in=60,min distance=0.1cm,#5](mid3),
(mid1) --[#2, ultra thick,#6](mid3),

@@

\newcommand{\scaleIntCtune}[4]{ {
\begin{tikzpicture}[baseline=-3]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {};
\vertex (a2) at (-.8, 0.6) {};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex [dot,scale=3](mid5) at (2,0){};
\vertex [dot,scale=2.5,hgrey0](mid6) at (2,0){};
\vertex (a3) at (2.8, .6) {};
\vertex (a4) at (2.8, -.6) {};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[#1, ultra thick,out=60,in=120,min distance=0.4cm,#2](mid3),
(mid3) --[#1, ultra thick,in=-60,out=-120,min distance=0.4cm,#2](mid1),
(mid3) --[#3, ultra thick,out=60,in=120,min distance=0.4cm,#4](mid5),
(mid5) --[#3, ultra thick,in=-60,out=-120,min distance=0.4cm,#4](mid3),
(mid5) --[photon, ultra thick](a4),
(mid5) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\title{Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence}

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  For the first time, we exploit special properties of these even-point theories to generalize the direct extraction of Feynman integral coefficients using $D$-dimensional unitarity cuts to the multi-loop case. In doing so, we find that the leading IR divergence of NLSM amplitudes exponentiates. We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms required to restore $U(1)$ behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  We also demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator in the $D$-dimensional description of Born-Infeld theory. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violates $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) superfields. Finally we find that $\mathcal{N}=4$ DBIVA amplitudes permit double-copy construction through two-loop order. This is shown by reproducing our unitarity based result with a loop-level double copy between $\mathcal{N}=4$ super-Yang-Mills numerators and our two-loop NLSM amplitudes. This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

\section{Introduction}\label{sec:intro}
Recent decades have witnessed a renaissance in our ability to compute scattering amplitudes to high orders in perturbation theory. At the epicenter of this explosion in $S$-matrix data is the unitarity method \cite{UnitarityMethod, Fusing, BDKUniarityReview}, which bypasses the standard Feynman diagram approach by constructing amplitudes directly from on-shell information. At one-loop, the unitarity method allows one to extract amplitudes by directly computing a series of unitarity cuts \cite{Forde:2007mi}. On-shell methods have furthermore unveiled novel amplitude-level structure, like the color-kinematics duality \cite{BCJ} and the double-copy construction \cite{BCJLoop}, which are completely obscured by a Lagrangian-level description of the physics. While much of the research in perturbative calculations thus far has focused on gauge theory and gravity, recent literature has investigated to what extent the $S$-matrix of effective field theory is constrained by on-shell data \cite{Adams:2006sv,Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd,Carrasco:2019yyn,Arkani-Hamed:2020blm,Carrasco:2021ptp,Chi:2021mio,Bonnefoy:2021qgu,Carrasco:2022lbm,Carrasco:2022sck,Green:2022slj,Pavao:2022kog,Chen:2022shl,Chen:2023dcx,Brown:2023srz}. In this work, we take a quantum leap to the multi-loop sector of effective field theory. As we embark on this journey, we anticipate that further probes of perturbative effective field theory data will unravel a rich tapestry of discoveries, adding to the current knowledge base and paving the way for future advancements in the field.

@@

By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed} . In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 

@@

\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu + \varphi \partial_\mu \varphi)

@@

The action above in \eqn{DBIVALag} can be similarly re-expressed in terms of our canonical spacetime coordinates, giving us the following flat-space Lagrangian for maximally supersymmetric Born-Infeld theory in $D=10$: 
 \label{dbivaLag}
    \alpha'^{\,2}\mathcal{L}^{\text{DBIVA}}s = \sqrt{\text{det} \left(\eta_{\mu\nu}- \alpha' F_{\mu\nu} + \alpha'^{\,2} (\bar{\lambda} \Gamma_\mu \partial_\nu \lambda)- \alpha'^4(\bar{\lambda} \Gamma^\rho \partial_\mu \lambda )(\bar{\lambda} \Gamma_\rho \partial_\nu \lambda) \right)}
  \,.
To recover the $D=4$ theory, we simply dimensionally reduce in a way that preserves the maximum number of supercharges, which is $\mathcal{N}=4$ in $D=4$. While this is obscured at the level of the Lagrangian, implementing a dimensional reduction that preserves the supercharges of \eqn{dbivaLag} is rather straight forward in an on-shell framework. 
\mathcal{M}^{\text{DBIVA}} = A^{\text{NLSM}} \otimes A^{\text{sYM}}
Since the double copy construction holds $D$-dimensionally at tree-level \cite{BCJreview}, double-copying maximal sYM with NLSM will yield the analogous spectrum of maximally supersymmetry DBIVA theory, whether in $D=10$, as described above, or in $D=4$, as we will study in this work. 
\paragraph{Abelian Open String}
Similar to the amplitudes of DBIVA theories described above, the tree-level amplitudes for the open superstring (OSS) ~\cite{Mafra:2011nv,Mafra:2011nw} likewise permit a double copy construction \cite{Broedel:2013tta}. To construct open-superstring amplitudes we simply double copy maximally supersymmetric super-Yang-Mills (sYM) and $Z$-theory amplitudes~\cite{Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc}, as follows,
A^{\text{OSS}}= Z \otimes A^{ \text{sYM}}
where $\otimes$ implements the field theory double copy \cite{KLT, BCJ} described above. Moreover, the amplitudes of DBIVA pop out of the above construction as the field theory limit of the \textit{abelian open superstring} \cite{Green:1982sw}. To recover the abelian sector, one simply sums over all color orderings on the side of the bi-colored $Z$-theory amplitudes that obey string monodromy relations. In the field theory limit, the observables for so-called \textit{abelian} $Z$-theory are simply the amplitudes generated by the NLSM Lagrangian \cite{Carrasco:2016ldy}. That is, given a $n$-point amplitude in $Z$-theory, with field theory ordering $a=(a_1,a_2,...,a_n)$ and string theoretic color ordering $A=(A_1,A_2,...,A_n)$, one finds the following relation:
A^{\text{NLSM}}{(a_1,a_2,...,a_n)} \equiv \lim_{\alpha' \to 0} (\alpha')^{2-n}  Z_{\times}{(a_1,a_2,...,a_n)}
where $Z_{\times}$ are abelian $Z$-theory amplitudes of \cite{Carrasco:2016ldy}, defined by summing all possible orderings of Chan-Paton factors:
Z_{\times}(a_1,a_2,...,a_n)\equiv \sum_{A \in S^{n-1}} Z_{(A_1,...,A_{n-1},n)}{(a_1,a_2,...,a_n)}
Since the field theory limit of abelian $Z$-theory produces NLSM amplitudes, the field theory limit of the abelian OSS, gives rise to DBIVA observables at leading order:
\begin{eBox}
\lim_{\alpha' \to 0 }Z^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{\text{BAS}}\equiv A^{\text{NLSM}}
\Rightarrow \quad\lim_{\alpha' \to 0 }A^{\text{OSS}}_{\times}&= A^{\text{NLSM}} \otimes A^{ \text{sYM}}\equiv A^{\text{DBIVA}}
\end{eBox}
Additional details on the bi-colored $Z$-theory amplitudes these string-theoretic double-copy constructions can be found in \cite{Broedel:2013tta,Carrasco:2016ldy,Carrasco:2016ygv,Mafra:2016mcc,Azevedo:2018dgo} and references therein. We have boxed the two classes of amplitudes above, NLSM and DBIVA, to emphasize that these are the two theories that we will focus on in this work. However, the methods that we develop in the next section apply much more broadly to the full catalog of even-point theories. One could

@@

\begin{eBox}

@@

@@

@@

\end{eBox}

@@

where the $D$-dimensional charge conjugation operator can be defined in terms of the gamma matrices, $\Gamma_\mu$, as follows:

@@

@@
